---
title: "Pstat175 Final Project"
author: |
  | Students: 
  |  Stephanie Or  (3119294) 
  |  Jianpeng Yuan (7531445)
  |  Kesey Scoot   (3930955)
  |
  | Group Number: 23 
  |
  | Instructor: Adam Tashman
  
date: "December 6, 2019"
output: pdf_document
---

\clearpage

\tableofcontents
\clearpage

# Abstract

# Data Source and Backgriund information

```{r }  
# echo = FALSE
library(survival)
library(KMsurv)
# library(dplyr)
# library(ggplot2)
# library(GGally)
# library(rms)

data(pneumon)
head(pneumon,3)
dim(pneumon)
```
This data frame contains the following columns:  
chldage     -   Age child had pneumonia, months  
hospital    -  Indicator for hospitalization for pneumonia (1=yes, 0=no)  
mthage      -  Age of the mother, years  
urban       -  Urban environment for mother (1=yes, 0=no)  
alcohol     -  Alcohol use by mother during pregnancy (1=yes, 0=no)  
smoke       -  Cigarette use by mother during pregnancy (1=yes, 0=no)  
region      -  Region of the coutry (1=northeast, 2=north central, 3=south, 4=west)  
poverty     -  Mother at poverty level (1=yes, 0=no)  
bweight     -  Normal birthweight (>5.5 lbs.) (1=yes, 0=no)  
race        -  Race of the mother (1=white, 2=black, 3=other)  
education   -  Education of the mother, years of school  
nsibs       -  Number of siblings of the child  
wmonth      -  Month the child was weaned  
sfmonth     -  Month the child on solid food  
agepn       -  Age child in the hospital for pneumonia, months  

# Research Question

# Data Exploration
```{r}

summary(pneumon)

mean(pneumon$chldage) #mean of Age child had pneumonia in 
mean(pneumon$agepn)
length(which(pneumon$hospital=="1")) 
length(which(pneumon$hospital=="0")) 
#number of child Indicator for hospitalization for pneumonia

pneumon.fit <- survfit(Surv(pneumon$chldage,pneumon$hospital)~1)

plot(pneumon.fit,mark=19,lwd=2,ylim = c(0.3,1.0),
     col=c("steelblue","pink","pink"), 
     main="Kaplan–Meier estimator of the data")

summary(pneumon.fit)
print(pneumon.fit)

# why is it not working? All NA 
quantile(pneumon.fit, probs=c(.75,.50,.25),
         conf.int=FALSE) 
```


# Up-sample & Down-sample and Kaplan-Meier estimation curves
since we have 3397 censored and only 73 event in the original dataset, we are going to do up-sample and down-sample to get a better dataset. 
```{r}
table(pneumon$hospital)

set.seed(99)
balance_data <- function(df, method, dsize){
  event <- df[df$hospital=="1",]
  censored <- df[df$hospital=="0",]
  nevent <- nrow(event)
  ncensored <- nrow(censored)
  
  if(method == "down"){
    if(nevent > ncensored)
    {
      dfe <- events[sample(1:nevent, dsize, replace=F),]
      new_dataset <- rbind(censored,dfe)
    }
    else{ #nevent <= ncensored
      dfc <- censored[sample(1:ncensored, dsize, replace = F),]
      new_dataset <- rbind(event,dfc)
    }
    new_dataset
  }
  
  else if(method =="up"){
    if(nevent < ncensored){
      dfe <- event[sample(1:nevent, dsize, replace = T),]
      new_dataset <- rbind(censored,dfe)
    }
    else{ #nevent <= ncensored
      dfc <- censored[sample(1:ncensored, dsize, replace = T),]
      new_dataset <- rbind(event,dfc)
    }
  }
  new_dataset
}

plotKM <- function(dataset){
  pneumon.fit <- survfit(Surv(dataset$chldage,dataset$hospital)~1)
  # print(summary(pneumon.fit))
  print(pneumon.fit)
  plot(pneumon.fit,mark=19,lwd=2,ylim = c(0.3,1.0),
       col=c("steelblue","pink","pink"))
  # pneumon.cox <- coxph((Surv(chldage,hospital)~.), data = dataset)
  # print(pneumon.cox)
}

#down sample to 73
new_dataset_down <- balance_data(pneumon,method="down",dsize = 73)
table(new_dataset_down$hospital)
plotKM(new_dataset_down)
title("Kaplan–Meier estimator of the downsample 73 data") 

#up sample to 3397
new_dataset_up <- balance_data(pneumon,method="up", dsize = 3397)
table(new_dataset_up$hospital)
plotKM(new_dataset_up)
title("Kaplan–Meier estimator of the upsample to 3397 data") 

#up sample for event and down sample for censored 300 each 
new_dataset300 <- balance_data(pneumon,method="up", dsize = 300)
new_dataset300 <- balance_data(new_dataset300,method="down",dsize = 300)
table(new_dataset300$hospital)
summary(new_dataset300)
plotKM(new_dataset300)
title("Kaplan–Meier estimator of the up-down-sample to 300 data")
``` 
We are going to use new_dataset300, which will have 300 data for both censored and event. 

# Model Building
## KM Curves (TESTING?)
```{r}
pneumon300 <- new_dataset300

library(My.stepwise)
# Stepwise Variable Selection Procedure for  
# Cox’s Proportional Hazards Model and Cox’s Model

pneumon.variable.list <- c("mthage","urban", "alcohol","smoke", "region","poverty", "bweight","race", "education", "nsibs", "wmonth","sfmonth","agepn")
My.stepwise.coxph(Time = "chldage", Status = "hospital", variable.list = pneumon.variable.list , data = pneumon300)

pneumon300.fit <- coxph(Surv(chldage,hospital)~nsibs + urban + region +
                            mthage + wmonth + alcohol + smoke + agepn, data = pneumon300)
anova(pneumon300.fit)


#   Confidence Intervals for the Coefficients

plot(pneumon.fit,lwd=2,col=c(2,4),
fun="cloglog",xlab="Time until Death \n in days",ylab="log(S(t))")



## Cox Proportional Hazards Model
# pneumon300.cox <- coxph((Surv(chldage,hospital)~.),data = pneumon300)
# summary(pneumon300.cox)
# pneumon300.cox_fit <- survfit(pneumon300.cox)
```
## Cox PH

# Model Checking 
## Hypothesis Testing 

# Residual tests

# PH Assumpting
## C-log-log plot 

# Interaction term ??? 

# Answer Question / Discussion

# Concluson



