---
title: "Pstat175 Final Project"
author: |
| Students: 
  |  Stephanie Or  (3119294) 
  |  Jianpeng Yuan (7531445)
  |  Kesey Scoot   (3930955)
  |
  | Group Number: 23 
  |
  | Instructor: Adam Tashman
  |
  
  date: "December 6, 2019"
output: pdf_document
---

\clearpage

\tableofcontents
\clearpage

# Abstract

# Data Source and Backgriund information

```{r }  
# echo = FALSE
library(survival)
library(KMsurv)
library(MASS)
library(dplyr)
# library(ggplot2)
# library(GGally)
# library(rms)

data(pneumon)
head(pneumon,3)
dim(pneumon)
```
This data frame contains the following columns:  
chldage     -  Age child had pneumonia, months  
hospital    -  Indicator for hospitalization for pneumonia (1=yes, 0=no)  
mthage      -  Age of the mother, years  
urban       -  Urban environment for mother (1=yes, 0=no)  
alcohol     -  Alcohol use by mother during pregnancy (1=yes, 0=no)  
smoke       -  Cigarette use by mother during pregnancy (1=yes, 0=no)  
region      -  Region of the coutry (1=northeast, 2=north central, 3=south, 4=west)  
poverty     -  Mother at poverty level (1=yes, 0=no)  
bweight     -  Normal birthweight (>5.5 lbs.) (1=yes, 0=no)  
race        -  Race of the mother (1=white, 2=black, 3=other)  
education   -  Education of the mother, years of school  
nsibs       -  Number of siblings of the child  
wmonth      -  Month the child was weaned  
sfmonth     -  Month the child on solid food  
agepn       -  Age child in the hospital for pneumonia, months  

# Research Question

# Data Exploration
```{r}
summary(pneumon)

mean(pneumon$chldage) #mean of Age child had pneumonia in 
mean(pneumon$agepn)
length(which(pneumon$hospital=="1")) 
length(which(pneumon$hospital=="0")) 
#number of child Indicator for hospitalization for pneumonia

pneumon.fit <- survfit(Surv(pneumon$chldage,pneumon$hospital)~1)

plot(pneumon.fit,mark=19,lwd=2,ylim = c(0.3,1.0),
     col=c("steelblue","pink","pink"), 
     main="Kaplan–Meier estimator of ",
     xlab = "Time (hospitalized) by month"
)

summary(pneumon.fit)
print(pneumon.fit)

# why is it not working? All NA 
quantile(pneumon.fit, probs=c(.75,.50,.25),
         conf.int=FALSE) 
```

# Up-sample & Down-sample and Kaplan-Meier estimation curves
since we have 3397 censored and only 73 event in the original dataset, we are going to do up-sample and down-sample to get a better dataset. 
```{r}
table(pneumon$hospital)

set.seed(99)
balance_data <- function(df, method, dsize){
  event <- df[df$hospital=="1",]
  censored <- df[df$hospital=="0",]
  nevent <- nrow(event)
  ncensored <- nrow(censored)
  
  if(method == "down"){
    if(nevent > ncensored)
    {
      dfe <- events[sample(1:nevent, dsize, replace=F),]
      new_dataset <- rbind(censored,dfe)
    }
    else{ #nevent <= ncensored
      dfc <- censored[sample(1:ncensored, dsize, replace = F),]
      new_dataset <- rbind(event,dfc)
    }
    new_dataset
  }
  
  else if(method =="up"){
    if(nevent < ncensored){
      dfe <- event[sample(1:nevent, dsize, replace = T),]
      new_dataset <- rbind(censored,dfe)
    }
    else{ #nevent <= ncensored
      dfc <- censored[sample(1:ncensored, dsize, replace = T),]
      new_dataset <- rbind(event,dfc)
    }
  }
  new_dataset
}

plotKM <- function(dataset){
  pneumon.fit <- survfit(Surv(dataset$chldage,dataset$hospital)~1)
  print(pneumon.fit)
  plot(pneumon.fit,mark=19,lwd=2,ylim = c(0.3,1.0),
       col=c("steelblue","pink","pink"))
}
``` 

```{r}
#down sample to 73
new_dataset_down <- balance_data(pneumon,method="down",dsize = 73)
table(new_dataset_down$hospital)
plotKM(new_dataset_down)
title("Kaplan–Meier estimator of the downsample to 73 data each") 

#up sample to 3397
new_dataset_up <- balance_data(pneumon,method="up", dsize = 3397)
table(new_dataset_up$hospital)
plotKM(new_dataset_up)
title("Kaplan–Meier estimator of the upsample to 3397 data each") 

#up sample for event and down sample for censored 300 each 
new_dataset300 <- balance_data(pneumon,method="up", dsize = 300)
new_dataset300 <- balance_data(new_dataset300,method="down",dsize = 300)
table(new_dataset300$hospital)
plotKM(new_dataset300)
title("Kaplan–Meier estimator of the up-down-sample to 300 data each")
```
We are going to use new_dataset300, which will have 300 data for both censored and event. 

# cleaning/ changing data 
```{r}
new_dataset300 <- new_dataset300 %>%
  group_by(alcohol) %>%
  mutate(alcohol_1=ifelse(alcohol ==0,0, 
                          ifelse(alcohol==1 , 1, 
                                 ifelse(alcohol==2 , 1,
                                        ifelse(alcohol==3 , 1 ,1)))))
table(new_dataset300$alcohol_1)

new_dataset300 <- new_dataset300 %>%
  group_by(smoke) %>%
  mutate(smoke_1=ifelse(smoke ==0,0, 
                        ifelse(smoke==1 , 1, 
                               ifelse(smoke==2 , 1, 1))))
table(new_dataset300$smoke_1)
```
Change alcohol to 1 if it is 2,3 or 4 \n
change smoke to 1 if it is 2 or 3 \n 

# Model cleaning  ## cox PH model 
```{r}
pneumon300 <- new_dataset300

library(My.stepwise)
# Stepwise Variable Selection Procedure for Cox’s Proportional Hazards Model and Cox’s Model


## before changeing alcohol and smoke
pneumon.variable.list <- c("mthage","urban", "alcohol","smoke", "region","poverty", 
                           "bweight","race", "education", "nsibs", "wmonth","sfmonth","agepn")
My.stepwise.coxph(Time = "chldage", Status = "hospital",data = pneumon300,
                  variable.list = pneumon.variable.list )
pneumon300.coxph <- coxph(Surv(chldage,hospital)~nsibs + urban + region + 
                            mthage + wmonth + alcohol + smoke, data = pneumon300)
summary(pneumon300.coxph)


## after changeing alcohol and smoke
pneumon_1.variable.list <- c("mthage","urban","alcohol_1","smoke_1", "region","poverty", 
                             "bweight","race", "education", "nsibs", "wmonth","sfmonth","agepn")
My.stepwise.coxph(Time = "chldage", Status = "hospital", data = pneumon300,
                  variable.list = pneumon_1.variable.list )
pneumon300_1.coxph <- coxph(Surv(chldage,hospital)~ nsibs + urban + region + 
                              mthage + smoke_1 + wmonth , data = pneumon300)
summary(pneumon300_1.coxph)

anova(pneumon300.coxph,pneumon300_1.coxph)
```
After we use My,stepwise.coxph(), we think what the $poverty,bweight,race,education$ and $sfmonth$ is not that we are interested.  
Talk about that agepn is not useful (same reason with lab D)


# Model Building
## Kaplan-Meier estimation Curves (TESTING?)  
```{r}
# par(mfrow=c(1,2))

## nsibs       -  Number of siblings of the child  
table(pneumon300$nsibs) 
nsibs.km <- plot(survfit(Surv(chldage,hospital)~nsibs,data = pneumon300),
                 main ="Number of siblings of the child KM curves",
                 xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:6)
legend("bottomleft",inset = 0,legend=c("0","1","2","3","4"),
       pch=15,col=2:6)


## urban       -  Urban environment for mother (1=yes, 0=no)  
table(pneumon300$urban)
urban.km <- plot(survfit(Surv(chldage,hospital)~urban,data = pneumon300),
                 main ="Urban environment for mother KM curves",
                 xlab="Survival Probability", ylab = "Time", wd=2, col = c(2,4))
legend("bottomleft",inset = 0,legend=c("1=yes", "0=no"),
       pch=15,col=c(2,4))


## region      -  Region of the coutry (1=northeast, 2=north central, 3=south, 4=west)  
table(pneumon300$region)
region.km <- plot(survfit(Surv(chldage,hospital)~region,data = pneumon300),
                  main ="Region of the coutry KM curves",
                  xlab="Survival Probability", ylab = "Time",lwd=2, col = 2:5)
legend("bottomleft",inset = 0,pch=15,col= 2:5,
       legend=c("1=northeast", "2=north central","3=south", "4=west"))


## mthage      -  Age of the mother, years  
table(pneumon300$mthage)
mthage <- plot(survfit(Surv(chldage,hospital)~mthage,data = pneumon300),
               main ="Age of the mother, years KM curves",
               xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:15)
legend("bottomleft",inset =0,pch=15,col= 2:15,
       legend=c("16","17","18","19","20","21","22","23","24","25","26","27","28","29"))


## wmonth      -  Month the child was weaned  
table(pneumon300$wmonth)
wmonth.km <- plot(survfit(Surv(chldage,hospital)~wmonth,data = pneumon300),
                  main ="Month the child was weaned KM curves",
                  xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:16)
legend("bottomleft",inset =0,pch=15,col= 2:16,
       legend=c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","22"))


## alcohol     -  Alcohol use by mother during pregnancy (1=yes, 0=no)
par(mfrow=c(1,2))
table(pneumon300$alcohol_1)
alcohol_1.km <- plot(survfit(Surv(chldage,hospital)~alcohol_1,data = pneumon300),
                  main ="Alcohol use by mother during pregnancy KM curves",
                  xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:3)
legend("bottomleft",inset =0,pch=15,col= 2:3, legend=c("0","1"))

table(pneumon300$alcohol)
alcohol.km <- plot(survfit(Surv(chldage,hospital)~alcohol,data = pneumon300),
                  main ="Alcohol use by mother during pregnancy KM curves",
                  xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:6)
legend("bottomleft",inset =0,pch=15,col= 2:6, legend=c("0","1","2","3","4"))


## smoke       -  Cigarette use by mother during pregnancy (1=yes, 0=no)
par(mfrow=c(1,2))
table(pneumon300$smoke_1) 
smoke_1.km <- plot(survfit(Surv(chldage,hospital)~smoke_1,data = pneumon300),
                 main ="Cigarette use by mother during pregnancy KM curves",
                 xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:3)
legend("bottomleft",inset =0,pch=15,col= 2:3, legend=c("0","1"))

## (1=yes, 0=no) 2=?
table(pneumon300$smoke) 
smoke.km <- plot(survfit(Surv(chldage,hospital)~smoke,data = pneumon300),
                 main ="Cigarette use by mother during pregnancy KM curves",
                 xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:4)
legend("bottomleft",inset =0,pch=15,col= 2:4, legend=c("0","1","2"))


## agepn       -  Age child in the hospital for pneumonia, months  
par(mfrow=c(1,1))
table(pneumon300$agepn) #0:12
agepn.km <- plot(survfit(Surv(chldage,hospital)~agepn,data = pneumon300),
                 main ="Age child in the hospital for pneumonia in months KM curves",
                 xlab="Survival Probability", ylab = "Time",lwd=2, col = 2:13)
legend("bottomleft",inset =0,pch=15,col= 2:13,
       legend=c("0","1","2","3","4","5","6","7","8","9","10","11","12"))
```



#Log rank test
```{r}
survdiff(Surv(chldage,hospital)~nsibs,data = pneumon300)

survdiff(Surv(chldage,hospital)~urban,data = pneumon300)

survdiff(Surv(chldage,hospital)~region,data = pneumon300)

survdiff(Surv(chldage,hospital)~mthage,data = pneumon300)

survdiff(Surv(chldage,hospital)~wmonth,data = pneumon300)

survdiff(Surv(chldage,hospital)~alcohol_1,data = pneumon300) #P-value= 1 > 0.05
survdiff(Surv(chldage,hospital)~alcohol,data = pneumon300)   #P-value=0.6 > 0.05

survdiff(Surv(chldage,hospital)~smoke_1,data = pneumon300)
survdiff(Surv(chldage,hospital)~smoke,data = pneumon300)
```

## AIC
```{r}
stepAIC(pneumon300.coxph, direction = "backward")

#after change the dataset for alcohol and smoke
stepAIC(pneumon300_1.coxph, direction = "backward")
```

## Residual tests
```{r}
cox.zph(pneumon300.coxph)  ##?alcohol too small?

#after change the dataset for alcohol and smoke
cox.zph(pneumon300_1.coxph) ##
```
if before changing alcohol = 0.0098 < 0.05 too small? 
As a result, we decide to stratify the variable alcohol. 

if after changing smoke_1 = 0.0024 < 0.05  too small? 
As a result, we decide to stratify the variable smoke_1.  

## C-log-log plot
```{r}
## nsibs       -  Number of siblings of the child  
nsibs.km <- plot(survfit(Surv(chldage,hospital)~nsibs,data = pneumon300),
                 main ="c-log-log plot for nsibs variable", fun="cloglog",
                 xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:6)
legend("topleft",inset = 0,legend=c("0","1","2","3","4"),
       pch=15,col=2:6)


## urban       -  Urban environment for mother (1=yes, 0=no)  
urban.km <- plot(survfit(Surv(chldage,hospital)~urban,data = pneumon300),
                 main ="c-log-log plot for urban variable",fun="cloglog",
                 xlab="Survival Probability", ylab = "Time", wd=2, col = c(2,4))
legend("topleft",inset = 0,legend=c("1=yes", "0=no"),
       pch=15,col=c(2,4))


## region      -  Region of the coutry (1=northeast, 2=north central, 3=south, 4=west)  
region.km <- plot(survfit(Surv(chldage,hospital)~region,data = pneumon300),
                  main ="c-log-log plot for region variable", fun="cloglog",
                  xlab="Survival Probability", ylab = "Time",lwd=2, col = 2:5)
legend("topleft",inset = 0,pch=15,col= 2:5,
       legend=c("1=northeast", "2=north central","3=south", "4=west"))


## mthage      -  Age of the mother, years  
mthage <- plot(survfit(Surv(chldage,hospital)~mthage,data = pneumon300),
               main ="c-log-log plot for mthage variable", fun="cloglog",
               xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:15)
legend("topleft",inset =0,pch=15,col= 2:15,
       legend=c("16","17","18","19","20","21","22","23","24","25","26","27","28","29"))


## wmonth      -  Month the child was weaned  
wmonth.km <- plot(survfit(Surv(chldage,hospital)~wmonth,data = pneumon300),
                  main ="c-log-log plot for wmonth variable", fun="cloglog",
                  xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:16)
legend("topleft",inset =0,pch=15,col= 2:16,
       legend=c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","22"))


## alcohol     -  Alcohol use by mother during pregnancy (1=yes, 0=no)
par(mfrow=c(1,2))
alcohol_1.km <- plot(survfit(Surv(chldage,hospital)~alcohol_1,data = pneumon300),
                  main ="c-log-log plot for alcohol_1 variable", fun="cloglog",
                  xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:3)
legend("topleft",inset =0,pch=15,col= 2:3, legend=c("0","1"))

alcohol.km <- plot(survfit(Surv(chldage,hospital)~alcohol,data = pneumon300),
                  main ="c-log-log plot for alcohol variable",fun="cloglog",
                  xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:6)
legend("topleft",inset =0,pch=15,col= 2:6, legend=c("0","1","2","3","4"))


## smoke       -  Cigarette use by mother during pregnancy (1=yes, 0=no)
par(mfrow=c(1,2))
smoke_1.km <- plot(survfit(Surv(chldage,hospital)~smoke_1,data = pneumon300),
                 main ="c-log-log plot for smoke_1 variable", fun="cloglog",
                 xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:3)
legend("topleft",inset =0,pch=15,col= 2:3, legend=c("0","1"))

## (1=yes, 0=no) 2=?
smoke.km <- plot(survfit(Surv(chldage,hospital)~smoke,data = pneumon300),
                 main ="c-log-log plot for smoke variable", fun="cloglog",
                 xlab="Survival Probability", ylab = "Time", lwd=2, col = 2:4)
legend("topleft",inset =0,pch=15,col= 2:3, legend=c("0","1","2"))

```

## Interaction term ??? 

## final model 

## Ensure that the PH Assumption is actually met

## Estimates with Confidence Intervals


# Answer Question / Discussion

# Concluson

# References 
Package 'My.stepwise' https://cran.r-project.org/web/packages/My.stepwise/My.stepwise.pdf
Package ‘KMsurv’ https://cran.r-project.org/web/packages/KMsurv/KMsurv.pdf
pneumon https://www.rdocumentation.org/packages/KMsurv/versions/0.1-5/topics/pneumon


# Appendix (All code) 


